---
title: "R package bp"
output: html_document
author: "Elias Ardila"
---

# Data

Load Dataset.

```{r setenv,message=FALSE,results='hide'}
Sys.setlocale("LC_TIME","en_US.UTF-8")
```
```{r loaddata,message=FALSE}
library(bp)
library(ggplot2)
library(rmarkdown)
library(kableExtra)
library(gridExtra)
```

```{r read data,message=FALSE,warnings=FALSE}
elias <- read.csv("elias.csv", sep=",")
elias$DATE <- as.Date(elias$DATE,format="%Y-%m-%d")
```

# Fitted Model
```{r create}
eliasbp <- process_data(elias,sbp='syst',dbp='diast', 
                        date_time='date.time',id='id',hr='hr')
```

# Plots

## LOESS  BP ~ Days

```{r plot1,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_ts_plots(eliasbp)[[1]]
png('images/datetime.png')
bp_ts_plots(eliasbp)[[1]]
dev.off()
```

## LOESS BP ~ Hour

```{r plot2,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_ts_plots(eliasbp)[[2]]
png('images/hour.png')
bp_ts_plots(eliasbp)[[2]]
dev.off()
```


## BP Classification by (Lee et al 2000)


   | BP Stage | Systolic(mmHg) | Diastolic(mmHg) |
   | ---------|:---------------:|-----------------:|
   |  Low (optional) |	<100  |	and 	<60   |
   |  Normal  |	<120 	| and 	<80 |
   |  Elevated |	120 - 129  |	and 	<80|
   |  Stage 1 - All |	130 - 139 |	and 	80 - 89|
   |  Stage 1 - ISH (ISH - S1) | 130 - 139 | 	and 	<80|
   |  Stage 1 - IDH (IDH - S1) |	<130 	|and 	80 - 89 |
   |  Stage 2 - All  |	>140  |	and 	>90|
   |  Stage 2 - ISH (ISH - S2)  |	>140 | 	and 	<90|
   |  Stage 2 - IDH (IDH - S2) |	<140 	| and 	>90 |
   |  Crisis (optional) |	>180 | 	or 	>120|


## All BP values

```{r plot4,message=FALSE,results='hide',echo=FALSE, fig.align='center'}
bp_hist(eliasbp)[[1]]
png('images/hist1.png')
bp_hist(eliasbp)[[1]]
dev.off()
```


## Systolic
    
```{r plot5,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_hist(eliasbp)[[2]]
png('images/hist2.png')
bp_hist(eliasbp)[[2]]
dev.off()
```

## Diastolic

```{r plot6,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_hist(eliasbp)[[3]]
png('images/hist3.png')
bp_hist(eliasbp)[[3]]
dev.off()
```

## Scatterplot BP  Systolic ~ Diastolic

```{r plot3,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_scatter(eliasbp)
png('images/scatter.png')
bp_scatter(eliasbp)
dev.off()
```

## Day of Week


```{r plotdayweek}
tage <- bp_tables(eliasbp)$CLASS_Day_of_Week
newnames <- c(rownames(tage),"Sum")
sumtage <- apply(tage[,1:8],2,sum)
rownames(tage) = NULL
tage <- rbind(tage,sumtage)
rownames(tage) <- newnames
tage%>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r daysbp, echo=FALSE,message=FALSE,results='hide'}
png("images/daysofweek.png",height=200,width=800)
xxx <- tableGrob(tage)
grid.arrange(xxx)
dev.off()
```

## Time of Day

```{r }
timeday <- bp_tables(eliasbp)$CLASS_Time_of_Day
newnames <- c(rownames(timeday),"Sum")
sumtimeday <- apply(timeday[,1:5],2,sum)
rownames(timeday) = NULL
timeday <- rbind(timeday,sumtimeday)
rownames(timeday) <- newnames
timeday %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r daysbpx, echo=FALSE,message=FALSE,results='hide'}
png("images/timeday.png",height=200,width=800)
xxx <- tableGrob(timeday)
grid.arrange(xxx)
dev.off()
```

##  American Heart Association Classification

| Blood Pressure Category |	Systolic (mmHg) 	|	Diastolic (mmHg) |
| ---------|:---------------:|-----------------:|                     
|Low (Hypotension) | 	Less than 100  |	and 	Less than 60 |
|Normal            |	100 - 120      |	and 	60 - 80      |
|Elevated 	   |    120 - 129      |	and 	60 - 80      |
|Stage 1 Hypertension | 	130 - 139  |	or 	80 - 89      |
|Stage 2 Hypertension |	140 - 180 	   |or 	90 - 120             |
|Hypertensive Crisis  |	Higher than 180    | 	and/or 	Higher than 120 |


```{r stages}
eliasstages <- bp_stages(elias,sbp='syst',dbp='diast',bp_type="ap")
dummysbp <- table(eliasstages$SBP_CATEGORY)
dummydbp <- table(eliasstages$DBP_CATEGORY)
```

```{r,message=FALSE,results='hide',echo=FALSE}
sbpstages <- c(dummysbp[2], dummysbp[1],dummysbp[3],dummysbp[4])
dbpstages <- dummydbp
```

## Systolic 

```{r barplot1, fig.align='center'}
df <- data.frame(dummysbp)
df_hist1 <- df
colnames(df) <- c("Class","Freq")
ggplot(data=df,aes(x=factor(Class,level=c("Normal","Elevated","Stage 1","Stage 2")),y=Freq,fill=Class)) +
       geom_bar(stat='identity',alpha=0.65,color='gray57')  +
       scale_fill_manual(values = c("Normal"="darkgreen","Elevated"="yellow","Stage 1"="orangered1","Stage 2"="red")) + xlab("Category") + guides(fill=guide_legend(title="SBP Category")) + theme(legend.position='none') + theme_minimal()
```

```{r plotapsbp, message=FALSE,results='hide',echo=FALSE}
##barplot(sbpstages,col=c("green","yellow","orange","red"))
##grid(nx=NA,ny=NULL)
png('images/sbpstages.png')
ggplot(data=df,aes(x=factor(Class,level=c("Normal","Elevated","Stage 1","Stage 2")),y=Freq,fill=Class)) +
       geom_bar(stat='identity',alpha=0.65, color='gray57')  +
       scale_fill_manual(values = c("Normal"="darkgreen","Elevated"="yellow","Stage 1"="orangered1","Stage 2"="red")) + xlab("Category") + guides(fill=guide_legend(title="SBP Category")) + theme(legend.position='none') + theme_minimal()
dev.off()
```

## Diastolic

```{r barplot2, fig.align='center' }
df2 <- data.frame(dummydbp)
colnames(df2) <- c("Class","Freq")
ggplot(data=df2,aes(x=Class,y=Freq,fill=Class)) +
       geom_bar(stat='identity',alpha=0.65, color='gray57')  +
       scale_fill_manual(values = c("Normal"="darkgreen","Elevated"="yellow","Stage 1"="orangered1","Stage 2"="red") ) + xlab("Category") + guides(fill=guide_legend(title="DBP Category")) + theme(legend.position='none') + theme_minimal()
```

```{r plotapdbp, message=FALSE,results='hide',echo=FALSE}
##barplot(dbpstages,col=c("green","orange","red"))
##grid(nx=NA,ny=NULL)
png('images/dbpstages.png')
ggplot(data=df2,aes(x=Class,y=Freq,fill=Class)) +
       geom_bar(stat='identity',alpha=0.65,color='gray57')  +
       scale_fill_manual(values = c("Normal"="darkgreen","Elevated"="yellow","Stage 1"="orangered1","Stage 2"="red")) + xlab("Category") + guides(fill=guide_legend(title="DBP Category")) + theme(legend.position='none') + theme_minimal()
dev.off()
```


# Blood Pressure Metrics

## Coefficient of Variation (Munter et al 2011)

```{r variation}
kv <- bp_cv(eliasbp)
kv[,-1] <- round(kv[,-1],2)
kv %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r variat, echo=FALSE,message=FALSE,results='hide'}
png("images/coeffvar.png",height=200,width=800)
xxx <- tableGrob(kv,rows=NULL)
grid.arrange(xxx)
dev.off()
```

## Average Real Variability (Mena et al 2005)

```{r avr}
wv <- bp_arv(eliasbp)
wv[,-1] <- round(wv[,-1],2)
wv %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r avvar, echo=FALSE,message=FALSE,results='hide'}
png("images/avvar.png",height=200,width=800)
xxx <- tableGrob(wv,rows=NULL)
grid.arrange(xxx)
dev.off()
```


## Successive Variation (Munter et al 2011)

```{r sucv}
scv <- bp_sv(eliasbp)
scv[,-1] <- round(scv[,-1],2)
scv %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r sucvar, echo=FALSE,message=FALSE,results='hide'}
png("images/sucvar.png",height=200,width=800)
xxx <- tableGrob(scv,rows=NULL)
grid.arrange(xxx)
dev.off()
```

## Blood Pressure Magnitude (peak and trough) (Munter et al 2011)

```{r presmag}
psm <- bp_mag(eliasbp)
psm[,-1] <- round(psm[,-1],2)
psm %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r pressmag, echo=FALSE,message=FALSE,results='hide'}
png("images/pressmag.png",height=200,width=800)
xxx <- tableGrob(psm,rows=NULL)
grid.arrange(xxx)
dev.off()
```

# Bayesian Playground

```{r bayesian,message=FALSE}
library(rstan)
n.iter <- 100000
n.chains <- 4
```

## Systolic

````{verbatim}
data{
int<lower=1> N;
real y[N];
real nu;
}
parameters{
real<lower=50,upper=200> mu;
real<lower=0> tau;
}
transformed parameters{
real<lower=0> sigma;
sigma = sqrt(1/tau);
}
model{
for(i in 1:N){
y[i] ~ student_t(nu,mu,sigma);
}
mu ~ uniform(50,200);
tau ~ gamma(0.001,0.001);
}
````

```{r fitstan, message=FALSE,results='hide'}
syst.dat <- list(N=length(elias$SYST),y=elias$SYST,nu=4)
syst.fit.stan <- stan(file="syst.stan",data=syst.dat,chains=n.chains,iter=n.iter)
```

```{r stanoutput}
print(syst.fit.stan)
```


```{r stanplot,fig.align='center'}

plot(syst.fit.stan,show_density=TRUE,pars=c("mu"))

```

```{r stanplotsyst,echo=FALSE,message=FALSE,results='hide'}
png("images/stanplotsyst.png",height=600,width=800)
plot(syst.fit.stan,show_density=TRUE,pars=c("mu"))
dev.off()

```

## Diastolic

````{verbatim}
data{
int<lower=1> N;
real y[N];
real nu;
}
parameters{
real<lower=60,upper=100> mu;
real<lower=0> tau;
}
transformed parameters{
real<lower=0> sigma;
sigma = sqrt(1/tau);
}
model{
for(i in 1:N){
y[i] ~ student_t(nu,mu,sigma);
}
mu ~ uniform(60,100);
tau ~ gamma(0.001,0.001);
}
````

```{r fitstan2, message=FALSE,results='hide'}
diast.dat <- list(N=length(elias$DIAST),y=elias$DIAST,nu=4)
diast.fit.stan <- stan(file="diast.stan",data=diast.dat,chains=n.chains,iter=n.iter)
```

```{r stanoutput2}
print(diast.fit.stan)
```

```{r stanplot2,fig.align='center'}

plot(diast.fit.stan,show_density=TRUE,pars=c("mu"))

```

```{r stanplotdiast,echo=FALSE,message=FALSE,results='hide'}
png("images/stanplotdiast.png",height=600,width=800)
plot(diast.fit.stan,show_density=TRUE,pars=c("mu"))
dev.off()
```

# Local Level Model

## Systolic

```{r dlm}
library(dlm)
```

```{r local_level_syst}

build <- function(parm){
    dlmModPoly(order=1,dV=parm[1],dW=parm[2])
}

syst <- elias$SYST

fit1 <- dlmMLE(syst, c(1.0,1.0),build)

dlm.syst.fitted <- build(fit1$par)

dlm.syst <- dlmFilter(syst, dlm.syst.fitted)

dlm.a <- dlm.syst$a
dlm.a[1] <- NA


dlm.syst.smooth <- dlmSmooth(syst, dlm.syst.fitted)
```

```{r locallevelsystfilter ,fig.align='center'}
plot(syst,pch=19,type="p",xlab="",ylab="")
lines(dlm.a)
```

```{r locallevelsystsmoothing, fig.align='center'}
plot(syst,pch=19,type="p",xlab="",ylab="")
lines(dlm.syst.smooth$s)

```

## Local Linear Model

```{r , message=FALSE }
library(KFAS)
```

```{r }
kfas_local_linear <- function(theta,daten,a0=c(0,0),P0=c(100,100)){
    ## =====================================
    ## SigmaEpsilon       |    SigmaNu     |
    ##      1             |       2        |
    ##      1             |      2:3       |
    ## =====================================
 
    Zt <- matrix(nrow=1,ncol=2,data=c(1,0))
    Zt
    Tt <- matrix(nrow=2,ncol=2,data=c(1,0,1,1))
    Tt
    Rt <- diag(2)
    Rt
    SigmaEpsilon <- matrix(nrow=1,ncol=1,theta[1])
    SigmaEpsilon
    SigmaNu <- diag(theta[2:3])
    SigmaNu
    
    a1 <- a0
    a0
    P1 <- diag(P0)
    P1
    ##P1inf <- matrix(nrow=1,ncol=1,1)
    mm <- SSModel(daten ~  -1 + SSMcustom(Z=Zt,
                                        T=Tt,
                                        Q=SigmaNu,
                                        R=Rt,
                                        a1=a1,
                                        P1=P1,
                                        ##P1inf=P1inf
                                        ),H=SigmaEpsilon)
    mod <- KFS(mm)
    return(mod)
}
```

```{r, echo=FALSE}
model_kfas_local_linear <- function(theta,daten,a0=c(0,0),P0=c(100,100)){
    ## =====================================
    ## SigmaEpsilon       |    SigmaNu     |
    ##      1             |       2        |
    ##      1             |      2:3       |
    ## =====================================
 
    Zt <- matrix(nrow=1,ncol=2,data=c(1,0))
    Zt
    Tt <- matrix(nrow=2,ncol=2,data=c(1,0,1,1))
    Tt
    Rt <- diag(2)
    Rt
    SigmaEpsilon <- matrix(nrow=1,ncol=1,theta[1])
    SigmaEpsilon
    SigmaNu <- diag(theta[2:3])
    SigmaNu
    
    a1 <- a0
    a0
    P1 <- diag(P0)
    P1
    ##P1inf <- matrix(nrow=1,ncol=1,1)
    mm <- SSModel(daten ~  -1 + SSMcustom(Z=Zt,
                                        T=Tt,
                                        Q=SigmaNu,
                                        R=Rt,
                                        a1=a1,
                                        P1=P1,
                                        ##P1inf=P1inf
                                        ),H=SigmaEpsilon)

    return(mm)
}
```

```{r }
llkfas_local_linear <- function(theta,daten,a0=c(0,0),P0=c(100,100)){
    mod <- kfas_local_linear(theta,daten,a0,P0)
    return(-mod$logLik)
}
```

## Systolic

### Filtered

```{r }
theta <- rep(0.01,3)
lower <- c(1e-6,1e-6,1e-6)
```

```{r optimlocallevel, message=FALSE, results='hide'}
fit1 <- optim(par=theta,fn=llkfas_local_linear,daten=syst,lower=lower,method="L-BFGS-B",control=list(trace=1))
```

```{r }
kfas.optim <- kfas_local_linear(fit1$par,syst)

kfas.linear.att <- kfas.optim$att
kfas.linear.att[1,1] <- NA 
V_mu <- as.vector(kfas.optim$V_mu)

N <- length(syst)

ul <- rep(0,N)
ll <- rep(0,N)
ulpt <- rep(0,N)
llpt <- rep(0,N)

i <- 1

while(i <= N){
   ulpt[i] <- kfas.linear.att[i,1] + qnorm(1-0.025)*sqrt(diag(kfas.optim$Ptt[,,i])[1])
   llpt[i] <- kfas.linear.att[i,1] + qnorm(0.025)*sqrt(diag(kfas.optim$Ptt[,,i])[1])
   ul[i] <- kfas.optim$alphahat[i,1] +  qnorm(1-0.025)*sqrt(V_mu[i])
   ll[i] <- kfas.optim$alphahat[i,1] + qnorm(0.025)*sqrt(V_mu[i])
   i <- i + 1
}
```

```{r, message=FALSE}
library(zoo)
```

```{r locallinearfilter, fig.align='center'}
plot(syst ,pch=19,type="o",xlab="",ylab="",col="darkgrey",ylim=c(100,160))
lines(kfas.linear.att[,1],lty=2)
lines(ulpt,lty=3,col="red")
lines(llpt,lty=3,col="red")
legend("bottomright",legend=c("Systolic","Filtered","95% limits"),lty=c(1,2,3),pch=c(19,NA,NA),
        col=c("darkgrey","black","red"),cex=.9,bty="n")
```


```{r kfaslocallinearfilter,echo=FALSE,message=FALSE,results='hide'}
png("images/locallinearfilter.png",height=600,width=800)
plot(syst ,pch=19,type="o",xlab="",ylab="",col="darkgrey",ylim=c(100,160))
lines(kfas.linear.att[,1],lty=2)
lines(ulpt,lty=3,col="red")
lines(llpt,lty=3,col="red")
legend("bottomright",legend=c("Systolic","Filtered","95% limits"),lty=c(1,2,3),pch=c(19,NA,NA),
        col=c("darkgrey","black","red"),cex=.9,bty="n")
dev.off()
```

```{r }
mod.kfas <- model_kfas_local_linear(fit1$par,syst)
syst.pred <- predict(mod.kfas,n.ahead=5)
```

```{r }
alphahat <- coredata(kfas.optim$alphahat)
nsyst <- dim(alphahat)[1]
xt <- 1:nsyst
n.ahead <- 5
xxt <- (nsyst+1):(nsyst+n.ahead)
k.alphahat <- apply(alphahat,1,sum)

```

## Smoothed State


```{r locallinearsmoothing, fig.align='center'}
plot(syst ~ xt,pch=19,type="o",xlab="",ylab="", xlim=c(1, nsyst + n.ahead),col="darkgrey",ylim=c(100,160))
lines(alphahat[,1] ~ xt,lty=2)
lines(coredata(syst.pred) ~ xxt,type="o",col="magenta",pch=19)
lines(ul,lty=3,col="red")
lines(ll,lty=3,col="red")
legend("bottomright",legend=c("Systolic","Smoothed","Predicted","95% limits" ),lty=c(1,2,1,1),pch=c(19,NA,19,NA),
col=c("darkgrey","black","magenta","red"),cex=.9,bty="n")
```

```{r locallinearsmoothingforecast,echo=FALSE,message=FALSE,results='hide'}
png("images/forecastkfas.png",height=600,width=800)
plot(syst ~ xt,pch=19,type="o",xlab="",ylab="", xlim=c(1, nsyst + n.ahead),col="darkgrey",ylim=c(100,160))
lines(alphahat[,1] ~ xt,lty=2)
lines(coredata(syst.pred) ~ xxt,type="o",col="magenta",pch=19)
lines(ul,lty=3,col="red")
lines(ll,lty=3,col="red")
legend("bottomright",legend=c("Systolic","Smoothed","Predicted","95% limits" ),lty=c(1,2,1,1),pch=c(19,NA,19,NA),
col=c("darkgrey","black","magenta","red"),cex=.9,bty="n")
dev.off()
```

## DLM

```{r }
library(dlm)
theta <- c(0.1,0.1,0.1)

funModel <- function(parm){
   m1 <- dlmModPoly(dV=exp(parm[1]),dW=exp(parm[2:3]))
   return(m1)
}

oMLE.DLM <- dlmMLE(syst, parm=log(theta),build=funModel)

stopifnot(oMLE.DLM$convergence == 0)

oFitted.DLM <- funModel(oMLE.DLM$par)

oFiltered.DLM <- dlmFilter(syst,oFitted.DLM)

oSmoothed.DLM <- dlmSmooth(syst,oFitted.DLM)

tsSmoState.DLM <- dropFirst(oSmoothed.DLM$s)

ff <- oFiltered.DLM$f
ff[1] <- NA

```

```{r,fig.align='center' }
plot(syst,pch=19)
lines(ff,col="red")
```

```{r, fig.align='center'}
plot(syst, pch=19)
lines(oSmoothed.DLM$s[,1],col="red")
legend("topright",legend=c("Systolic","smoothed state"), lty=c(1,1),col=c("black","red"))
```

## Forecast. n.Ahead=5

```{r }
f_f <- 5
fut1 <- dlmForecast(oFiltered.DLM,n=f_f)
n.syst <- length(syst)

tdx <- seq(n.syst + 1,n.syst + f_f)
```

```{r,fig.align='center' }
plot(syst,type="o",col="darkgrey",xlab="",xlim=c(0,tdx[f_f]),pch=19,ylab="",ylim=c(100,160))
lines(ff,lty=2)
lines(as.numeric(fut1$f) ~ tdx,type="o",pch=16,col="red")
legend("bottomright",legend=c("Systolic","Filtered","Predicted"),lty=c(1,2,1),pch=c(19,NA,19),
col=c("darkgrey","black","red"),cex=.9,bty="n")
```


```{r forecastn5,echo=FALSE,message=FALSE,results='hide'}
png("images/forecast.png",height=600,width=800)
plot(syst,type="o",col="darkgrey",xlab="",xlim=c(0,tdx[f_f]),pch=19,ylab="",ylim=c(100,160))
lines(ff,lty=2)
lines(as.numeric(fut1$f) ~ tdx,type="o",pch=16,col="red")
legend("bottomright",legend=c("Systolic","Filtered","Predicted"),lty=c(1,2,1),pch=c(19,NA,19),
col=c("darkgrey","black","red"),cex=.9,bty="n")
dev.off()
```

## Diastolic

### Filtered

```{r }
diast <- elias$DIAST
theta <- rep(0.01,3)
lower <- c(1e-6,1e-6,1e-6)
```

```{r optimlocallevel.diast, message=FALSE, results='hide'}
fit1.diast <- optim(par=theta,fn=llkfas_local_linear,daten=diast,lower=lower,method="L-BFGS-B",control=list(trace=1))
```

```{r }
kfas.optim.diast <- kfas_local_linear(fit1.diast$par,diast)

kfas.linear.att.diast <- kfas.optim.diast$att
kfas.linear.att.diast[1,1] <- NA 
V_mu.diast <- as.vector(kfas.optim.diast$V_mu)

N <- length(diast)

ul <- rep(0,N)
ll <- rep(0,N)
ulpt.diast <- rep(0,N)
llpt.diast <- rep(0,N)

i <- 1

while(i <= N){
   ulpt.diast[i] <- kfas.linear.att.diast[i,1] + qnorm(1-0.025)*sqrt(diag(kfas.optim.diast$Ptt[,,i])[1])
   llpt.diast[i] <- kfas.linear.att.diast[i,1] + qnorm(0.025)*sqrt(diag(kfas.optim.diast$Ptt[,,i])[1])
   ul[i] <- kfas.optim.diast$alphahat[i,1] +  qnorm(1-0.025)*sqrt(V_mu.diast[i])
   ll[i] <- kfas.optim.diast$alphahat[i,1] +  qnorm(0.025)*sqrt(V_mu.diast[i])

   i <- i + 1
}


```

```{r, message=FALSE}
library(zoo)
```

```{r locallinearfilter.diast , fig.align='center'}
plot(diast ,pch=19,type="o",xlab="",ylab="",col="darkgrey",ylim=c(60,100))
lines(kfas.linear.att.diast[,1],lty=2)
lines(ulpt.diast , lty=3,col="red")
lines(llpt.diast , lty=3,col="red")
legend("bottomright",legend=c("Diastolic","Filtered","95% limits"),lty=c(1,2,3),pch=c(19,NA,NA),
        col=c("darkgrey","black","red"),cex=.9,bty="n")
```


```{r kfaslocallinearfilter.diast,echo=FALSE,message=FALSE,results='hide'}
png("images/locallinearfilterdiast.png",height=600,width=800)
plot(diast ,pch=19,type="o",xlab="",ylab="",col="darkgrey",ylim=c(60,100))
lines(kfas.linear.att.diast[,1],lty=2)
lines(ulpt.diast , lty=3,col="red")
lines(llpt.diast , lty=3,col="red")
legend("bottomright",legend=c("Diastolic","Filtered","95% limits"),lty=c(1,2,3),pch=c(19,NA,NA),
        col=c("darkgrey","black","red"),cex=.9,bty="n")
dev.off()
```

```{r }
mod.kfas.diast <- model_kfas_local_linear(fit1.diast$par,diast)
diast.pred <- predict(mod.kfas.diast,n.ahead=5)
```

```{r }
alphahat <- coredata(kfas.optim.diast$alphahat)
nsyst <- dim(alphahat)[1]
xt <- 1:nsyst
n.ahead <- 5
xxt <- (nsyst+1):(nsyst+n.ahead)
k.alphahat <- apply(alphahat,1,sum)

```

## Smoothed State


```{r locallinearsmoothing.diast, fig.align='center'}
plot(diast ~ xt,pch=19,type="o",xlab="",ylab="", xlim=c(1, nsyst + n.ahead),col="darkgrey",ylim=c(60,100))
lines(alphahat[,1] ~ xt,lty=2)
lines(coredata(diast.pred) ~ xxt,type="o",col="magenta",pch=19)
lines(ul,lty=3,col="red")
lines(ll,lty=3,col="red")
legend("bottomright",legend=c("Diastolic","Smoothed","Predicted","95% limits"),lty=c(1,2,1,1),pch=c(19,NA,19,NA),
col=c("darkgrey","black","magenta","red"),cex=.9,bty="n")
```

```{r locallinearsmoothingforecast.diast,echo=FALSE,message=FALSE,results='hide'}
png("images/forecastkfasdiast.png",height=600,width=800)
plot(diast ~ xt,pch=19,type="o",xlab="",ylab="", xlim=c(1, nsyst + n.ahead),col="darkgrey",ylim=c(60,100))
lines(alphahat[,1] ~ xt,lty=2)
lines(coredata(diast.pred) ~ xxt,type="o",col="magenta",pch=19)
lines(ul,lty=3,col="red")
lines(ll,lty=3,col="red")
legend("bottomright",legend=c("Diastolic","Smoothed","Predicted","95% limits"),lty=c(1,2,1,1),pch=c(19,NA,19,NA),
col=c("darkgrey","black","magenta","red"),cex=.9,bty="n")
dev.off()
```

## DLM

```{r }
library(dlm)
theta <- c(0.1,0.1,0.1)

funModel <- function(parm){
   m1 <- dlmModPoly(dV=exp(parm[1]),dW=exp(parm[2:3]))
   return(m1)
}

oMLE.DLM <- dlmMLE(diast, parm=log(theta),build=funModel)

stopifnot(oMLE.DLM$convergence == 0)

oFitted.DLM <- funModel(oMLE.DLM$par)

oFiltered.DLM <- dlmFilter(diast,oFitted.DLM)

oSmoothed.DLM <- dlmSmooth(diast,oFitted.DLM)

tsSmoState.DLM <- dropFirst(oSmoothed.DLM$s)

ff <- oFiltered.DLM$f
ff[1] <- NA

```

```{r,fig.align='center' }
plot(diast,pch=19)
lines(ff,col="red")
```

```{r, fig.align='center'}
plot(diast, pch=19)
lines(oSmoothed.DLM$s[,1],col="red")
legend("topright",legend=c("Systolic","smoothed state"), lty=c(1,1),col=c("black","red"))
```

## Forecast. n.Ahead=5

```{r }
f_f <- 5
fut1 <- dlmForecast(oFiltered.DLM,n=f_f)
n.diast <- length(diast)

tdx <- seq(n.diast + 1,n.diast + f_f)
```

```{r,fig.align='center' }
plot(diast,type="o",col="darkgrey",xlab="",xlim=c(0,tdx[f_f]),pch=19,ylab="",ylim=c(60,100))
lines(ff,lty=2)
lines(as.numeric(fut1$f) ~ tdx,type="o",pch=16,col="red")
legend("bottomright",legend=c("Diastolic","Filtered","Predicted"),lty=c(1,2,1),pch=c(19,NA,19),
col=c("darkgrey","black","red"),cex=.9,bty="n")
```


```{r forecastn5.diast,echo=FALSE,message=FALSE,results='hide'}
png("images/forecastdiast.png",height=600,width=800)
plot(diast,type="o",col="darkgrey",xlab="",xlim=c(0,tdx[f_f]),pch=19,ylab="",ylim=c(60,100))
lines(ff,lty=2)
lines(as.numeric(fut1$f) ~ tdx,type="o",pch=16,col="red")
legend("bottomright",legend=c("Diastolic","Filtered","Predicted"),lty=c(1,2,1),pch=c(19,NA,19),
col=c("darkgrey","black","red"),cex=.9,bty="n")
dev.off()
```
