---
title: "R package bp"
output: html_document
author: "Elias Ardila"
---

# Data

Load Dataset.

```{r setenv,message=FALSE,results='hide'}
Sys.setlocale("LC_TIME","en_US.UTF-8")
```
```{r loaddata,message=FALSE}
library(bp)
library(ggplot2)
library(rmarkdown)
library(kableExtra)
##library(xtable)
library(gridExtra)
```

```{r read data,message=FALSE,warnings=FALSE}
elias <- read.csv("elias.csv", sep=",")
elias$DATE <- as.Date(elias$DATE,format="%Y-%m-%d")
```

# Fitted Model
```{r create}
eliasbp <- process_data(elias,sbp='syst',dbp='diast', 
                        date_time='date.time',id='id',hr='hr')
```

# Plots

## LOESS  BP ~ Days

```{r plot1,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_ts_plots(eliasbp)[[1]]
png('images/datetime.png')
bp_ts_plots(eliasbp)[[1]]
dev.off()
```

## LOESS BP ~ Hour

```{r plot2,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_ts_plots(eliasbp)[[2]]
png('images/hour.png')
bp_ts_plots(eliasbp)[[2]]
dev.off()
```


## BP Classification by (Lee et al 2000)


   | BP Stage | Systolic(mmHg) | Diastolic(mmHg) |
   | ---------|:---------------:|-----------------:|
   |  Low (optional) |	<100  |	and 	<60   |
   |  Normal  |	<120 	| and 	<80 |
   |  Elevated |	120 - 129  |	and 	<80|
   |  Stage 1 - All |	130 - 139 |	and 	80 - 89|
   |  Stage 1 - ISH (ISH - S1) | 130 - 139 | 	and 	<80|
   |  Stage 1 - IDH (IDH - S1) |	<130 	|and 	80 - 89 |
   |  Stage 2 - All  |	>140  |	and 	>90|
   |  Stage 2 - ISH (ISH - S2)  |	>140 | 	and 	<90|
   |  Stage 2 - IDH (IDH - S2) |	<140 	| and 	>90 |
   |  Crisis (optional) |	>180 | 	or 	>120|


## All BP values

```{r plot4,message=FALSE,results='hide',echo=FALSE, fig.align='center'}
bp_hist(eliasbp)[[1]]
png('images/hist1.png')
bp_hist(eliasbp)[[1]]
dev.off()
```


## Systolic
    
```{r plot5,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_hist(eliasbp)[[2]]
png('images/hist2.png')
bp_hist(eliasbp)[[2]]
dev.off()
```

## Diastolic

```{r plot6,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_hist(eliasbp)[[3]]
png('images/hist3.png')
bp_hist(eliasbp)[[3]]
dev.off()
```

## Scatterplot BP  Systolic ~ Diastolic

```{r plot3,message=FALSE,results='hide',echo=FALSE,fig.align='center'}
bp_scatter(eliasbp)
png('images/scatter.png')
bp_scatter(eliasbp)
dev.off()
```

## Day of Week


```{r plotdayweek}
tage <- bp_tables(eliasbp)$CLASS_Day_of_Week
newnames <- c(rownames(tage),"Sum")
sumtage <- apply(tage[,1:8],2,sum)
rownames(tage) = NULL
tage <- rbind(tage,sumtage)
rownames(tage) <- newnames
tage%>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r daysbp, echo=FALSE,message=FALSE,results='hide'}
png("images/daysofweek.png",height=200,width=800)
xxx <- tableGrob(tage)
grid.arrange(xxx)
dev.off()
```

## Time of Day

```{r }
timeday <- bp_tables(eliasbp)$CLASS_Time_of_Day
newnames <- c(rownames(timeday),"Sum")
sumtimeday <- apply(timeday[,1:5],2,sum)
rownames(timeday) = NULL
timeday <- rbind(timeday,sumtimeday)
rownames(timeday) <- newnames
timeday %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r daysbpx, echo=FALSE,message=FALSE,results='hide'}
png("images/timeday.png",height=200,width=800)
xxx <- tableGrob(timeday)
grid.arrange(xxx)
dev.off()
```

##  American Heart Association Classification

| Blood Pressure Category |	Systolic (mmHg) 	|	Diastolic (mmHg) |
| ---------|:---------------:|-----------------:|                     
|Low (Hypotension) | 	Less than 100  |	and 	Less than 60 |
|Normal            |	100 - 120      |	and 	60 - 80      |
|Elevated 	   |    120 - 129      |	and 	60 - 80      |
|Stage 1 Hypertension | 	130 - 139  |	or 	80 - 89      |
|Stage 2 Hypertension |	140 - 180 	   |or 	90 - 120             |
|Hypertensive Crisis  |	Higher than 180    | 	and/or 	Higher than 120 |


```{r stages}
eliasstages <- bp_stages(elias,sbp='syst',dbp='diast',bp_type="ap")
dummysbp <- table(eliasstages$SBP_CATEGORY)
dummydbp <- table(eliasstages$DBP_CATEGORY)
```

```{r,message=FALSE,results='hide',echo=FALSE}
sbpstages <- c(dummysbp[2], dummysbp[1],dummysbp[3],dummysbp[4])
dbpstages <- dummydbp
```

## Systolic 

```{r barplot1, fig.align='center'}
df <- data.frame(dummysbp)
colnames(df) <- c("Class","Freq")
ggplot(data=df,aes(x=factor(Class,level=c("Normal","Elevated","Stage 1","Stage 2")),y=Freq,fill=Class)) +
       geom_bar(stat='identity',alpha=0.65,color='gray57')  +
       scale_fill_manual(values = c("Normal"="darkgreen","Elevated"="yellow","Stage 1"="orangered1","Stage 2"="red")) + xlab("Category") + guides(fill=guide_legend(title="SBP Category")) + theme(legend.position='none') + theme_minimal()
```

```{r plotapsbp, message=FALSE,results='hide',echo=FALSE}
##barplot(sbpstages,col=c("green","yellow","orange","red"))
##grid(nx=NA,ny=NULL)
png('images/sbpstages.png')
ggplot(data=df,aes(x=factor(Class,level=c("Normal","Elevated","Stage 1","Stage 2")),y=Freq,fill=Class)) +
       geom_bar(stat='identity',alpha=0.65, color='gray57')  +
       scale_fill_manual(values = c("Normal"="darkgreen","Elevated"="yellow","Stage 1"="orangered1","Stage 2"="red")) + xlab("Category") + guides(fill=guide_legend(title="SBP Category")) + theme(legend.position='none') + theme_minimal()
dev.off()
```

## Diastolic

```{r barplot2, fig.align='center' }
df2 <- data.frame(dummydbp)
colnames(df2) <- c("Class","Freq")
ggplot(data=df2,aes(x=Class,y=Freq,fill=Class)) +
       geom_bar(stat='identity',alpha=0.65, color='gray57')  +
       scale_fill_manual(values = c("Normal"="darkgreen","Elevated"="yellow","Stage 1"="orangered1","Stage 2"="red") ) + xlab("Category") + guides(fill=guide_legend(title="DBP Category")) + theme(legend.position='none') + theme_minimal()
```

```{r plotapdbp, message=FALSE,results='hide',echo=FALSE}
##barplot(dbpstages,col=c("green","orange","red"))
##grid(nx=NA,ny=NULL)
png('images/dbpstages.png')
ggplot(data=df2,aes(x=Class,y=Freq,fill=Class)) +
       geom_bar(stat='identity',alpha=0.65,color='gray57')  +
       scale_fill_manual(values = c("Normal"="darkgreen","Elevated"="yellow","Stage 1"="orangered1","Stage 2"="red")) + xlab("Category") + guides(fill=guide_legend(title="DBP Category")) + theme(legend.position='none') + theme_minimal()
dev.off()
```


# Blood Pressure Metrics

## Coefficient of Variation (Munter et al 2011)

```{r variation}
kv <- bp_cv(eliasbp)
kv[,-1] <- round(kv[,-1],2)
kv %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r variat, echo=FALSE,message=FALSE,results='hide'}
png("images/coeffvar.png",height=200,width=800)
xxx <- tableGrob(kv,rows=NULL)
grid.arrange(xxx)
dev.off()
```

## Average Real Variability (Mena et al 2005)

```{r avr}
wv <- bp_arv(eliasbp)
wv[,-1] <- round(wv[,-1],2)
wv %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r avvar, echo=FALSE,message=FALSE,results='hide'}
png("images/avvar.png",height=200,width=800)
xxx <- tableGrob(wv,rows=NULL)
grid.arrange(xxx)
dev.off()
```


## Successive Variation (Munter et al 2011)

```{r sucv}
scv <- bp_sv(eliasbp)
scv[,-1] <- round(scv[,-1],2)
scv %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r sucvar, echo=FALSE,message=FALSE,results='hide'}
png("images/sucvar.png",height=200,width=800)
xxx <- tableGrob(scv,rows=NULL)
grid.arrange(xxx)
dev.off()
```

## Blood Pressure Magnitude (peak and trough) (Munter et al 2011)

```{r presmag}
psm <- bp_mag(eliasbp)
psm[,-1] <- round(psm[,-1],2)
psm %>% kbl() %>% kable_paper("hover",full_width=FALSE)
```

```{r pressmag, echo=FALSE,message=FALSE,results='hide'}
png("images/pressmag.png",height=200,width=800)
xxx <- tableGrob(psm,rows=NULL)
grid.arrange(xxx)
dev.off()
```
